---
title: 'Lab 7: Clustering (ungraded)'
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidycensus)
library(tidyverse)
library(knitr)
library(sf)
library(tmap)
```

## Clustering

In this short ungraded lab, you will be downloading data from U.S. Census for Georgia census tracts and using hierarchical clustering to identify areas with similar characteristics across the state. This type of process is commonly used to classify regions or identify market segmentation.

Let's start by downloading census data from the 2016-2020 American Community Survey. You'll need to have already installed the U.S. Census API to do so. In this case, I'm going to use median household income for several racial classifications, commuting times (travel time to work), % with a Bachelor's degree, access to broadband, and median age. 

Feel free to play around with this code to choose different variables if you like--just remember to transform population counts to rates. It's not terribly hard to find other variables you might like to include.

```{r}
#Get full variable list
v20 <- load_variables(2020, "acs5", cache = TRUE)

#Identify relevant tables for our variables of interest and select them from the variable list
tables<-c("B15003","B08012","B01002","B19013","B28011")

var_select<-v20 %>%
  filter(substr(name,1,6) %in% tables)

#Download data and pivot wider
ga_tracts<-get_acs(geography="tract",variables=var_select$name,state="GA",year=2020)

ga_tracts_wide<-ga_tracts %>%
  select(-moe) %>%
  pivot_wider(names_from=variable,values_from=estimate)

#Select and rename our key variables
#You will need to play with this code if you download different variables
ga_tracts_sample<-ga_tracts_wide %>%
  rename(medage=B01002_001,
         medinc=B19013_001) %>%
  mutate(ba_pct=B15003_022/B15003_001*100,
         trav_10_20=(B08012_004+B08012_005)/B08012_001*100,
         trav_20_30=(B08012_006+B08012_007)/B08012_001*100,
         trav_30more=(B08012_008+B08012_009+B08012_010+B08012_011+B08012_012+B08012_013)/B08012_001*100,
         broadband=B28011_004/B28011_001*100,
         nointernet=B28011_008/B28011_001*100) %>%
  select(GEOID,NAME,medage,medinc,ba_pct:trav_30more,broadband,nointernet)

#Remove any rows with NAs
ga_tracts_sample<-na.omit(ga_tracts_sample)
```

We'll also load the centroids of census tracts using tidycensus. We'll load just the count of total population.

```{r}
tct_centroids<-get_acs(geography="tract",state="GA",var="B01001_001",geometry=TRUE,year=2020) %>%
  st_centroid() %>%
  inner_join(ga_tracts_sample %>% select("GEOID"))
```


###Part 1: Importing data and descriptive statistics###

Now that we have a dataset create clusters from these variables using HCA

**Question 1** _Subset the above data to select only the variables you want to use in this analysis. You should at least remove the geographic name fields so only the continuous variables remain. Choose as few as three variables or up to all of them (though that will be harder to interpret). Call the head of the table with kable when you are done._

```{r}
#Code here
```

**Question 2** _Create a distance matrix for the variables you selects using the `dist` function and then plot a dendogram from that matrix. Identify the number of clusters you would like to use in this analysis and create a bounding box using `rect.hclust`. Explain why you chose the number of clusters you did._

```{r}
#Code here
```

{Justification here.}

**Question 3** _Use the cutree function to add a list of cluster identifiers back to the `ga_tracts_sample` data frame. Call the head of the table using kable when you are done._

```{r}
#Code here
```

**Question 4** _Use a join to connect the cluster table you created to the tract centroids we have downloaded. Map the clusters from this data frame using sf. Summarise any spatial patterns you see._

```{r}
#Code here
```

{Summary here.}

**Question 5** _Using ggplot, create boxplots or some similar measure of distribution to examine the differences in values across your clusters._

```{r}
#Code here
```

**Question 6** _Based on these boxplots, give a name to each of your clusters. Briefly explain each one_

{Discussion here.}