---
title: 'Geog6300: Lab 6'
output: github_document
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

## Regression

**Value:** 30 points

```{r setup, message=FALSE, warning=FALSE}
library(sf)
library(tidyverse)
```

**Overview:**
This lab focuses on regression techniques. You'll be analyzing the association of county demographics on the presidential vote in 2012 (the Obama-Romney election).

###Part 1: Importing data and descriptive statistics###
This lab uses the same county data file you used in Lab 6. You also have a county shapefile that we'll use for mapping these results.

**Question 1** _Load the two files in the data folder--the election data and county level shapefile. Filter the election data for just the 2012 election, then join it to the county shapefile. Call the head of this data frame using kable when you are done._

```{r}
#Code goes here.
```

For this analysis, you'll be looking at the following variables:

* gop_pct: The % of vote for Romney (the Republican) in 2012
* medinc: Median household income
* wht_pop_pct: % of population classified as white
* ed_graddeg_pct: % of the population whose highest educational attainment is a graduate degree.

You'll be evaluating how each of the last three variables was associated with the county's Republican vote in 2012. To start with, you should assess the distribution of each variable.

**Question 2** _Create a QQ plot for each of the four variables of interest. Based on these graphs, assess the normality of these variables._

```{r}
#Code goes here.
```

{Your assessment of the variables' normality goes here.}

**Question 3** _Use tmap to map the four variables of interest using Jenks natural breaks as the classification method For an extra, optional challenge, you can dissolve the counties to state boundaries using group_by (with the state variable)and summarise. Add these boundaries to your maps and reduce the opacity of the county boundaries using the border.alpha parameter._

```{r}

```


**Question 4** _Based on the maps from question 3, summarise two major patterns you see in the spatial distribution of these data from any of your variables of interest. Be sure to make specific references to your data to illustrate your points._

{Your response here.}


**Question 5** _Create univariate regression models for each of the independent variables (income, % white, and % w/graduate degree) using % voting GOP as the dependent variable. Make sure the results of each model are visible._

```{r}
#Code goes here.
```

**Question 6** _Summarize the results of the models in question 5, focusing on the direction and significance of the model coefficients, the normality of model residuals, and the overall R2 value._

**Question 7** _Create a multivariate regression model with all three variables of interest and call the results._

```{r}
#Code goes here
```

**Question 8** _Summarise the results of the multivariate model. What are the direction, magnitude, and significance of each coefficient, and how did it change from the univariate models (if at all)? What do the R2 and F-statistic values tell you about overall model fit?_

{Response goes here}

**Question 9** _Use a QQ plot and a map to assess the normality of residuals and any spatial autocorrelation. Summarise any notable patterns that you see._

```{r}
#Code goes here
```

{Summary goes here}

**Question 10** _Assess any issues with multicollinearity or heteroskedastity in this model using the techniques shown in class. Run the appropriate tests and explain what their results show you._

```{r}
#Code goes here
```

{Interpretation goes here}

**Question 11** _Looking at the full model and your diagnostics, do you feel this is a model that provides meaningful results? Explain your answer._

**Lab reflection:** _How do you feel about the work you did on this lab? Was it easy, moderate, or hard? What were the biggest things you learned by completing it?_


**Bonus challenge**
There's lots of controversy over election maps, as choropleth versions can overemphaize land area compared to population. For this bonus challenge, map some aspect of these election results (for example, gop or dem vote, the margin between them, or any other variable you might create) using one of the following techniques:

* Cartograms: https://cran.r-project.org/web/packages/cartogram/readme/README.html
* Dot density maps: https://rpubs.com/mathetal/dotdensity
* Bivariate maps (maybe with a demographic variable): https://cran.r-project.org/web/packages/biscale/vignettes/biscale.html

Alternatively, create your own visualization of these election data using ggplot that isn't a choropleth map (or even a map at all). Post your map below and explain what you like about it.

```{r}
#Code goes here
```

{Interpretation goes here.}

